---
title: "Mini-Project 2"
author: "Your Name Here"
date: "October 31, 2017"
output:
  html_document:
    code_folding: hide
---


## Loading the data

This project is so interesting and relevant to my life!

```{r, include=FALSE}
load("house_elections.rda")
load("candidates.rda")
load("committees.rda")
load("contributions.rda")
```

```{r}
library(tidyverse)
library(gridExtra)
library(dplyr)
```

<<<<<<< HEAD
=======
```{r}
# Filtering out contributions that were against a candidate
negative_contributions <- contributions %>% 
  filter(transaction_type == "24A")

# Filtering out contributions that were for a candidate
positive_contributions <- contributions %>% 
  filter(transaction_type == "24E")

# Renaming a fec_id to cand_id for joining
h_elections <- house_elections %>% 
  rename(cand_id = fec_id)

# Finding candidates that were funded against
negative_candidates <- inner_join(negative_contributions, candidates, by = "cand_id") %>% 

  right_join(h_elections , by = "cand_id") %>% # Linking the candidates to their names, not just id number
  filter(!is.na(transaction_type)) %>% # Do not want NA transactions
  filter(!is.na(transaction_type)) %>% 
  group_by(cand_name.y, cand_party_affiliation, primary_votes,
           runoff_votes, general_votes, ge_winner) %>% 
  summarize(total_dollars = sum(transaction_amt)) %>% 
  arrange(desc(total_dollars))

positive_candidates <- inner_join(positive_contributions, candidates, by = "cand_id") %>% 
  right_join(h_elections, by = "cand_id") %>% 
  filter(!is.na(transaction_type)) %>% 
<<<<<<< HEAD
  group_by(cand_name.y, cand_party_affiliation, primary_votes,
           runoff_votes, general_votes, ge_winner) %>%
=======
  group_by(cand_name, cand_party_affiliation, primary_votes,
  runoff_votes, general_votes, ge_winner) %>%
>>>>>>> 052c36e75edb979fb545a4942d17e97b2b011b17
  summarize(total_dollars = sum(transaction_amt)) %>% 
  arrange(desc(total_dollars))
```


## Committees Dataset
tres_name: Treasurer's name
cmte_st1: Street 1 (Address)
cmte_st2: Street 2 (Address)
cmte_city: City or Town
cmte_state: State
cmte_zip: Zip Code
cmte_dsgn: Committee Designation
  + A = Authorized by a candidate
  + B = Lobbyist/Registrant PAC
  + D = Leadership PAC
  + J = Joint fundraiser
  + P = Principal campaign committee of a candidate
  + U = Unauthorized
cmte_type: Committee type
  + https://classic.fec.gov/finance/disclosure/metadata/CommitteeTypeCodes.shtml
cmte_party_affiliation: Party
  + https://classic.fec.gov/finance/disclosure/metadata/DataDictionaryPartyCodeDescriptions.shtml
cmte_filing_freq: Filing frequency
  + A = Administratively terminated
  + D = Debt
  + M = Monthly filer
  + Q = Quarterly filer
  + T = Terminated
  + W = Waived
org_type: Interest category group (seems important)
  + C = Corporation
  + L = Labor organization
  + M = Membership organization
  + T = Trade association
  + V = Cooperative
  + W = Corporation without capital stock
  
## Candidates Dataset
cand_id: Candidate identification.
  + A 9-character alpha-numeric code assigned to a candidate by the Federal Election Commission. The candidate ID for a specific candidate remains the same across election cycles as long as the candidate is running for the same office.
cand_name: Name of the candidate.
cand_party_affiliation: Party affiliation
  + https://classic.fec.gov/finance/disclosure/metadata/DataDictionaryPartyCodeDescriptions.shtml
cand_election_yr
>>>>>>> 0c2e86e24f3cc495c764016c07a153fb77739086
